---
title: "Introduction"
author: "Nicole Ortogero"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5,
  fig.height = 4,
  dpi=200
)
```


## Introduction

The NanoStringExperiment class was inherited from the SummarizedExperiment 
class. It was designed to encapsulate data and corresponding methods for data 
generated from NanoString platforms. This class serves as an intermediary 
class between SummarizedExperiment and NanoString platform-specific classes 
such as the NanoStringGeoMxSet class from the GeomxTools package. This 
package is meant to congregate common structures and methods to be used for 
extension to NanoString platform-specific packages.


There are numerous functions that NanoStringExperiment inherited from the 
SummarizedExperiment class. It is recommended to learn the 
SummarizedExperiment class prior to working with NanoStringExperiment objects.
You can find these in this link: 
https://bioconductor.org/packages/release/bioc/html/SummarizedExperiment.html

## Loading Packages

Loading the NanoStringExperiment package gives access to 
corresponding class and methods.

```{r import, eval=FALSE}
library(NanoStringExperiment)
```

## Building a NanoStringExperiment object from NanoString data 

Create a new NanoStringExperiment object using an 
expression matrix (assayData),
phenoData (AnnotatedDataFrame containing sample phenotypic metadata), 
protocolData (AnnotatedDataFrame containing sample protocol metadata), 
featureData (AnnotatedDataFrame containing feature metadata), 
experimentData (list or MIAME object containing experiment details), 
annotation (list containing annotation file names used), 
dimLabels (list of variables used as row and column keys in object).

```{r define}
datadir <- system.file("data", package="NanoStringExperiment")

demoExperiment <- NanoStringExperiment(
    assayData = readRDS(file.path(datadir, "testAssayData.rds")),
    phenoData = readRDS(file.path(datadir, "testPhenoData.rds")),
    protocolData = readRDS(file.path(datadir, "testProtocolData.rds")),
    featureData = readRDS(file.path(datadir, "testFeatureData.rds")),
    experimentData = readRDS(file.path(datadir, "testExperimentData.rds")),
    annotation = readRDS(file.path(datadir, "testAnnotation.rds")),
    dimLabels = readRDS(file.path(datadir, "testDimLabels.rds")))
demoExperiment
```

## Slots and accessors

The NanoStringExperiment class inherits the slots and accessors available
with any SummarizedExperiment class. There are a few added slots for data 
commonly used with NanoString experiments. Accessor and replacement methods 
have been added for these new slots.

```{r access}
assays(demoExperiment)
assay(demoExperiment, "exprs")[1:2, 1:2]
assay(demoExperiment, "exprsShift") <- assay(demoExperiment, "exprs") + 1
assay(demoExperiment, "exprsShift")[1:2, 1:2]

head(colData(demoExperiment))[1:2, ]
head(rowData(demoExperiment))[1:2, ]

#Version of package when object created
classVersion(demoExperiment)

annotation(demoExperiment)

design(demoExperiment) <- ~ `cell_line`
design(demoExperiment)
```

## Subsetting

Bracket notation and the subset method are two methods for subsetting 
NanoStringExperiment objects. This package also includes custom NanoString
subset functions such as to subset be negative features.

```{r subset}
demoExperiment[1:5, ]
demoExperiment[colData(demoExperiment)[["cell_line"]] == "HS578T", ]

subset(demoExperiment, 
    subset = CodeClass == "Negative", 
    select = cell_line == "HS578T")

negativeControlSubset(demoExperiment)
endogenousSubset(demoExperiment)
```


## Apply and summarize

As with a SummarizedExperiment object, you can apply functions across
assay data matrices using the apply function. The NanoStringExperiment 
package adds additional apply and summarize functions to facilitate working
with the object.

```{r applying}
head(assayDataApply(demoExperiment, MARGIN = 2, FUN = mean))

head(esBy(demoExperiment, 
            GROUP = "cell_line", 
            FUN = function(x) { 
              assayDataApply(x, MARGIN = 1, FUN = mean, elt = "exprs") 
            }))

summary(demoExperiment, MARGIN = 2, GROUP = "cell_line")[1:2]

with(demoExperiment, table(CodeClass))

head(munge(demoExperiment, mapping = ~`cell_line` + GeneMatrix))
```

## About this session
```{r session}
sessionInfo()
```
